The importance of serve in table tennis

### 2024-08-13 inst/table_tennis_serve.Rmd
magrittr package now needed explicitly
pair3()

```{r,label=defineH,message=FALSE}
library("hyper2",quietly=TRUE)
(H <- hyper2(pnames=c("S","a","b","c")))
```

so the likelihood function is uniform, for the only term is raised to
the zeroth power (players' names are not capitalised in R idiom for
clarity).  Now we have to input the data, and we will go through the
four scorelines above, one by one.  Note that this approach implicitly
assumes independence of scores.

### A vs B, A serves, score 5-1:
```{r,label=abaserve}
H[c("a","S")]     %<>% "+"(5  ) # a+S win 5 games
H["b"]            %<>% "+"(  1) # b wins 1 game
H[c("a","b","S")] %<>% "-"(5+1) # a+b+S play 1+5=6 games
H
```

@@

### A vs B, B serves, score 1-3:
```{r,label=abbserve}
H[c("b","S")]     %<>% "+"(3  ) # b+S win 3 games
H["a"]            %<>% "+"(  1) # a wins 1 game 
H[c("a","b","S")] %<>% "-"(3+1) # a+b+S play 1+3=4 games
```

### A vs C, A serves, score 4-1:
```{r,label=acaserve}
H[c("a","S")]     %<>% "+"(4  ) # a+S win 4 games
H["c"]            %<>% "+"(  1) # c wins 1 game 
H[c("a","c","S")] %<>% "-"(4+1) # a+c+S play 1+4=5 games
```

### A vs C, C serves, score 1-2:
```{r,label=accserve}
H["a"]            %<>% "+"(1  ) # a wins 1 game
H[c("c","S")]     %<>% "+"(  2) # c+S win 2 games
H[c("a","c","S")] %<>% "-"(1+2) # a+c+S play 1+2=3 games
```

## Likelihood function

@@

## Assessment of B vs C

@@

@@

>>>>

```{r,label=defineH,message=FALSE}
library("hyper2",quietly=TRUE)
library("magrittr")
(H <- hyper2(pnames=c("S","a","b","c")))
```

so the likelihood function is uniform.  We have players `a`, `b`, `c`
(players' names are not capitalised in R idiom for clarity) and a
reified entity `S` representing the strength of the serve.  Now we
have to input the data, and we will go through the four scorelines
above, one by one.  Note that this approach implicitly assumes
independence of scores.

### A vs B, A serves, score 5-1: {-}
```{r,label=abaserve}
H[c("a","S")]     %<>% "+"(5  ) # a+S win 5 games
H["b"]            %<>% "+"(  1) # b wins 1 game
H[c("a","b","S")] %<>% "-"(5+1) # a+b+S play 1+5=6 games
H
```

@@

* A vs B, B serves, score 1-3:
```{r,label=abbserve}
H[c("b","S")]     %<>% "+"(3  ) # b+S win 3 games
H["a"]            %<>% "+"(  1) # a wins 1 game 
H[c("a","b","S")] %<>% "-"(3+1) # a+b+S play 1+3=4 games
```

* A vs C, A serves, score 4-1:
```{r,label=acaserve}
H[c("a","S")]     %<>% "+"(4  ) # a+S win 4 games
H["c"]            %<>% "+"(  1) # c wins 1 game 
H[c("a","c","S")] %<>% "-"(4+1) # a+c+S play 1+4=5 games
```

* A vs C, C serves, score 1-2:
```{r,label=accserve}
H["a"]            %<>% "+"(1  ) # a wins 1 game
H[c("c","S")]     %<>% "+"(  2) # c+S win 2 games
H[c("a","c","S")] %<>% "-"(1+2) # a+c+S play 1+2=3 games
```

@@

```{r, label=equalplayerstrength,cache=TRUE}
samep.test(table_tennis,c("a","b","c"))
```

Above, `samep.test()` complains because it is performing a
one-dimensional optimization (we may ignore this issue).  The result
of the test is that there is no evidence to suggest a difference
between players `a`, `b`, and `c`.

@@

# Assessment of B vs C

@@

# Weighted likelihoods

We now try `hyper3` formalism:

```{r,label=calcproflike,cache=TRUE}
M_home <- matrix( c(NA, 5 , 4, 3, NA, 0, 2,0,NA),byrow=TRUE,3,3)
jj <- letters[1:3]
dimnames(M_home) <- list(jj,jj)

M_away <- matrix( c(NA, 1 , 1, 1, NA, 0, 1,0,NA),byrow=TRUE,3,3)
jj <- letters[1:3]
dimnames(M_away) <- list(jj,jj)
M_home
M_away

home_away3(M_home,M_away,lambda=1.8)

f <- function(lam){
    maxp(home_away3(M_home,M_away,lambda=lam),give=TRUE)$value
}
```

```{r,makelamlike,cache=TRUE}
lam <- exp(seq(from=log(0.6),to=log(20),len=9))
L <- sapply(lam,f)
```

```{r,optimlamlike,cache=TRUE}
maxf <- optimize(f,c(2,5),maximum=TRUE)
maxf
```

```{r plotlikelam}
L <- L-max(L)
plot(lam,L,type='b',pch=16,log="x")
abline(h=c(0,-2))
```

How does the situation change if we have some more observations?
Suppose `c` plays at home to `d`, and loses.  If we use
$\hat{\lambda}\simeq 3.35$:

```{r, label=newdatacd,cache=TRUE}
(Hnew <- home_away3(M_home,M_away,lambda=3.35) + pair3(c=0,d=1,lambda=3.35))
maxp(Hnew)
```

Above we see that our _estimate_ would be that `d` has a BT strength
of 1 and the other players have a strength of zero.  Is there strong
evidence that the teams' strength differ?

```{r,label=newdatatest,cache=TRUE}
equalp.test(Hnew)
```

Above we see that there is no evidence that the teams' strengths do in
fact differ.
